Investigating Cross-Sectional Size Distribution of Randomly Distributed 3D Spheres

### 2020-05-24 Paper/main.tex
added some citations

% \bibliography{example_bib.bib}

@@

Mathematically, determining the distribution of spherical particles from planar
sections is a well known problem called Wicksell's Corpuscle Problem
\todo{cite}, named after S. D. Wicksell, who solved the problem in 1925. With
this solution, it is possible to determine the original distribution of
spherical particles. Thus, it is possible to determmine the distribution of
particle sizes in a solid from its observed cross-section, allowing us to
identify and classify materials better \todo{nasa}.

@@

In order to obtain the planar cross-section, we employed the spherical cap
method \todo{cite}. If we consider a sphere with centre
$\overrightarrow{\textbf{c}_0} = (x_0, y_0, z_0)$ and radius $R$, and a plane
with equation $Ax + By + Cz = D$ such that $\overrightarrow{\textbf{n}} = (A, B,
C)$ is the normal vector to the plane, then the distance from the centre of the
circle to the closest point on the plane $p_0$ is given by $\rho$:
\begin{equation*}
\rho = \frac{(\overrightarrow{\textbf{c}_0} - \overrightarrow{\textbf{p}_0}) \cdot{}
  \overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{n}}|} = \frac{Ax_0 +
  By_0 + Cz_0 - D}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}
If $|\rho| < R$, then there is an intersection of the plane and the sphere. The
radius of the circle $r$ in the planar cross-section is given by:
\begin{equation*}
r = \sqrt{R^2 - \rho^2}
\end{equation*}
And we then binned this radius to produce a histogram of the distribution of
planar circle radii. It is also possible to recover the centre
$\overrightarrow{\textbf{c}}$ of the circle visible on the plane using:
\begin{equation*}
\overrightarrow{\textbf{c}} = \overrightarrow{\textbf{c}_0} +
\rho\frac{\overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{c}}|} = (x_0,
y_0, z_0) + \rho\frac{(A, B, C)}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}

@@

If $f(R)dR$ is the number of spheres per unit volume with radii in interval $[R,
R + dR]$, then we can define a distribution $\phi(x)dx$ which represents the
number of circles per unit area on the plane with radii in interval $[x,
x+dx]$ \todo{cite}. This distribution is given by:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{xf(R)}{\sqrt{R^2 - x^2}}
\label{eq:phi_OG}
\end{equation}
We can compute this integral to obtain a theoretical planar distribution for
different starting spherical radius distributions.

@@

If we compute the integral in Equation~\ref{eq:phi_OG}, replacing $f(R)$ with
$\delta(R-R_0)$ we find:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{x\delta(R-R_0)}{\sqrt{R^2 - x^2}} =
\frac{xH(R_0-x)}{\sqrt{R_0^2-x^2}}
\label{eq:phi_constant}
\end{equation}
where $H(R_0 - x)$ is the heaviside function, which ``turns on'' the
distribution only for $x < R0$ as it is not possible to have a circular radius
$x$ larger than the spherical radius. So we aim to see our experimental data
reproduce this distribution \todo{cite}. We also investigated how changing the
value of $R_0$ affects the experimental distribution.

@@

Indeed, we found that making the sphere sizes much larger made the placing of
the spheres in a finite space more difficult, and exponentially increased the
time complexity of the placing algorithm. This is because if the placed spheres
are too large, while it is initially easy to place spheres, as more are placed,
more coordinates must be randomly drawn before one if found for the next sphere
to be placed into. It is also possible that the algorithm never finds a possible
coordinate for later spheres. One possible solution to this may be to use a
better placing algorithm. Choosing a tight packing algorithm \todo{cite} would
maximise the number of spheres of a given radius that can be placed. However,
this would likely result in a less than random distribution of spheres if the
spheres are all of identical sizes. This would likely cause some systematic
error in the data, and so it was not investigated here. If we randomised the
sizes of the spheres as we do in the next part of the investigation, the problem
become significantly more complicated. Tight packing algorithms for irregular
shaped objects are very difficult to create and computationally expensive
\todo{cite}.

@@

However, it would be more prudent to use a Gaussian distribution to draw the
values of the noise. This is because it is likely that the many sources of error
possible in an experimental measurement of particle cross-sections in a material
will be convoluted together into a Gaussian distribution of errors due to the
central limit theorem \todo{cite}.

>>>>

\bibliography{ParticleSizeDistribution.bib}

@@

Mathematically, determining the distribution of spherical particles from planar
sections is a well known problem called Wicksell's Corpuscle Problem
\cite{DWicksell}, named after S. D. Wicksell, who solved the problem in 1925. With
this solution, it is possible to determine the original distribution of
spherical particles. Thus, it is possible to determmine the distribution of
particle sizes in a solid from its observed cross-section, allowing us to
identify and classify materials better \cite{Cuzzi2016}.

@@

In order to obtain the planar cross-section, we employed the spherical cap
method \cite{Sykes1922}. If we consider a sphere with centre
$\overrightarrow{\textbf{c}_0} = (x_0, y_0, z_0)$ and radius $R$, and a plane
with equation $Ax + By + Cz = D$ such that $\overrightarrow{\textbf{n}} = (A, B,
C)$ is the normal vector to the plane, then the distance from the centre of the
circle to the closest point on the plane $p_0$ is given by $\rho$:
\begin{equation*}
\rho = \frac{(\overrightarrow{\textbf{c}_0} - \overrightarrow{\textbf{p}_0}) \cdot{}
  \overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{n}}|} = \frac{Ax_0 +
  By_0 + Cz_0 - D}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}
If $|\rho| < R$, then there is an intersection of the plane and the sphere. The
radius of the circle $r$ in the planar cross-section is given by:
\begin{equation*}
r = \sqrt{R^2 - \rho^2}
\end{equation*}
And we then binned this radius to produce a histogram of the distribution of
planar circle radii. It is also possible to recover the centre
$\overrightarrow{\textbf{c}}$ of the circle visible on the plane using:
\begin{equation*}
\overrightarrow{\textbf{c}} = \overrightarrow{\textbf{c}_0} +
\rho\frac{\overrightarrow{\textbf{n}}}{|\overrightarrow{\textbf{c}}|} = (x_0,
y_0, z_0) + \rho\frac{(A, B, C)}{\sqrt{A^2 + B^2 + C^2}}
\end{equation*}

@@

If $f(R)dR$ is the number of spheres per unit volume with radii in interval $[R,
R + dR]$, then we can define a distribution $\phi(x)dx$ which represents the
number of circles per unit area on the plane with radii in interval $[x,
x+dx]$ \cite{Kiderlen2011}. This distribution is given by:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{xf(R)}{\sqrt{R^2 - x^2}}
\label{eq:phi_OG}
\end{equation}
We can compute this integral to obtain a theoretical planar distribution for
different starting spherical radius distributions.

@@

If we compute the integral in Equation~\ref{eq:phi_OG}, replacing $f(R)$ with
$\delta(R-R_0)$ we find:
\begin{equation}
\phi(x) = \int_{x}^{\infty}\frac{x\delta(R-R_0)}{\sqrt{R^2 - x^2}} =
\frac{xH(R_0-x)}{\sqrt{R_0^2-x^2}}
\label{eq:phi_constant}
\end{equation}
where $H(R_0 - x)$ is the heaviside function, which ``turns on'' the
distribution only for $x < R0$ as it is not possible to have a circular radius
$x$ larger than the spherical radius. So we aim to see our experimental data
reproduce this distribution. We also investigated how changing the
value of $R_0$ affects the experimental distribution.

@@

Indeed, we found that making the sphere sizes much larger made the placing of
the spheres in a finite space more difficult, and exponentially increased the
time complexity of the placing algorithm. This is because if the placed spheres
are too large, while it is initially easy to place spheres, as more are placed,
more coordinates must be randomly drawn before one if found for the next sphere
to be placed into. It is also possible that the algorithm never finds a possible
coordinate for later spheres. One possible solution to this may be to use a
better placing algorithm. Choosing a tight packing algorithm %\cite{Chen2001} would
maximise the number of spheres of a given radius that can be placed. However,
this would likely result in a less than random distribution of spheres if the
spheres are all of identical sizes. This would likely cause some systematic
error in the data, and so it was not investigated here. If we randomised the
sizes of the spheres as we do in the next part of the investigation, the problem
become significantly more complicated. Tight packing algorithms for irregular
shaped objects are very difficult to create and computationally expensive
\todo{cite}.

@@

However, it would be more prudent to use a Gaussian distribution to draw the
values of the noise. This is because it is likely that the many sources of error
possible in an experimental measurement of particle cross-sections in a material
will be convoluted together into a Gaussian distribution of errors due to the
central limit theorem \cite{Islam2018}.
